---
title: "RxNorm"
output:
  pdf_document: default
  html_document: default
  word_document: default

lang: en
output:
  Gmisc::docx_document:
    fig_caption: true
---

### Source: <https://rxnav.nlm.nih.gov/RxNormAPIs.html>

| TTY  |               Name                |                                                                                               Example                                                                                                | Description                                                                                                                                                                                        |
|:------------:|:------------:|:--------------------:|:----------------------|
|  IN  |            Ingredient             |                                                                                              Fluoxetine                                                                                              | A compound or moiety that gives the drug its distinctive clinical properties. Ingredients generally use the United States Adopted Name (USAN).                                                     |
| PIN  |        Precise Ingredient         |                                                                                       Fluoxetine Hydrochloride                                                                                       | A specified form of the ingredient that may or may not be clinically active. Most precise ingredients are salt or isomer forms                                                                     |
| MIN  |       Multiple Ingredients        |                                                                                       Fluoxetine / Olanzapine                                                                                        | Two or more ingredients appearing together in a single drug preparation, created from SCDF. In rare cases when IN/PIN or PIN/PIN combinations of the same base ingredient exist, created from SCD. |
| SCDC | Semantic Clinical Drug Component  |                                                                                          Fluoxetine 4 MG/ML                                                                                          | Ingredient + Strength                                                                                                                                                                              |
| SCDF |    Semantic Clinical Drug Form    |                                                                                       Fluoxetine Oral Solution                                                                                       | Ingredient + Dose Form                                                                                                                                                                             |
| SCDG | Semantic Clinical Dose Form Group |                                                                                       Fluoxetine Oral Product                                                                                        | Ingredient + Dose Form Group                                                                                                                                                                       |
| SCD  |      Semantic Clinical Drug       |                                                                                   Fluoxetine 4 MG/ML Oral Solution                                                                                   | Ingredient + Strength + Dose Form                                                                                                                                                                  |
| GPCK |           Generic Pack            |                                                           {11 (varenicline 0.5 MG Oral Tablet) / 42 (varenicline 1 MG Oral Tablet) } Pack                                                            | {# (Ingredient + Strength + Dose Form) / \# (Ingredient + Strength + Dose Form)} Pack                                                                                                              |
|  BN  |            Brand Name             |                                                                                                Prozac                                                                                                | A proprietary name for a family of products containing a specific active ingredient.                                                                                                               |
| SBDC |  Semantic Branded Drug Component  |                                                                                     Fluoxetine 4 MG/ML [Prozac]                                                                                      | Ingredient + Strength + Brand Name                                                                                                                                                                 |
| SBDF |    Semantic Branded Drug Form     |                                                                                  Fluoxetine Oral Solution [Prozac]                                                                                   | Ingredient + Dose Form + Brand Name                                                                                                                                                                |
| SBDG | Semantic Branded Dose Form Group  |                                                                                             Prozac Pill                                                                                              | Brand Name + Dose Form Group                                                                                                                                                                       |
| SBD  |       Semantic Branded Drug       |                                                                              Fluoxetine 4 MG/ML Oral Solution [Prozac]                                                                               | Ingredient + Strength + Dose Form + Brand Name                                                                                                                                                     |
| BPCK |          Brand Name Pack          | {12 (Ethinyl Estradiol 0.035 MG / Norethindrone 0.5 MG Oral Tablet) / 9 (Ethinyl Estradiol 0.035 MG / Norethindrone 1 MG Oral Tablet) / 7 (Inert Ingredients 1 MG Oral Tablet) } Pack [Leena 28 Day] | {# (Ingredient Strength Dose Form) / \# (Ingredient Strength Dose Form)} Pack [Brand Name]                                                                                                         |
| PSN  |         Prescribable Name         |                                                                                          Leena 28 Day Pack                                                                                           | Synonym of another TTY, given for clarity and for display purposes in electronic prescribing applications. Only one PSN per concept.                                                               |
|  SY  |              Synonym              |                                                                                     Prozac 4 MG/ML Oral Solution                                                                                     | Synonym of another TTY, given for clarity.                                                                                                                                                         |
| TMSY |    Tall Man Lettering Synonym     |                                                                                FLUoxetine 10 MG Oral Capsule [PROzac]                                                                                | Tall Man Lettering synonym of another TTY, given to distinguish between commonly confused drugs.                                                                                                   |
|  DF  |             Dose Form             |                                                                                            Oral Solution                                                                                             | See Appendix 2 for a full list of Dose Forms.                                                                                                                                                      |
|  ET  |       Dose Form Entry Term        |                                                                                             Nasal Drops                                                                                              | Entry terms for dose forms (TTY=DF).                                                                                                                                                               |
| DFG  |          Dose Form Group          |                                                                                             Oral Liquid                                                                                              | See Appendix 3 for a full list of Dose Form Groups.                                                                                                                                                |

: RxNorm Term Types (TTY)

## Setup

Get directory

```{r setup, eval=FALSE, include=TRUE, encoding = "UTF-8"}
getwd()

# First specify the packages of interest
packages = c("jsonlite", "readxl","readr","curl","tidyverse","stringr","xlsx","xlsx",
             "XLConnect","bit64","sqldf", "haven","httr")


# Now load or install&load all
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

## NDC to Drug Name

Using RxNorm API to transform pre-defined NDC code into drug name.

```{r NDC_to_DrugName, eval=FALSE, include=TRUE}
# Input

# Example
# NDCpath <- "/Users/peilinhuang/Dropbox (UFL)/Lynn Huang/Rxmix/11DigitNDC.xlsx"
# Outputpath <- "/Users/peilinhuang/Dropbox (UFL)/Lynn Huang/Rxmix/NDC_Name_r.xlsx"

NDCpath <- "/Users/peilinhuang/Dropbox (UFL)/Analysts shared folder/OneFL/OneFL_Lynn/OneFL_Phar_NDC.xlsx"
Outputpath <- "/Users/peilinhuang/Dropbox (UFL)/Analysts shared folder/OneFL/OneFL_Lynn/OneFL_Phar_NDC.xlsx"

# Create function
NDC_to_DrugName <- function(NDC){
  # Find Rxcui By NDC
  path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui.json?idtype=NDC&id=", NDC)
  RxCUI <- content(GET(url = path))$idGroup$rxnormId[[1]]
  RxCUI
  # Get Names By RxCUI
  path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui/", RxCUI, "/allProperties.json?prop=names")
  Drug_Name <- content(GET(url = path))$propConceptGroup$propConcept[[1]][["propValue"]]
  Drug_Name <- ifelse(is.null(Drug_Name),"",Drug_Name)
  return(Drug_Name)
}
# Test the function
NDC_to_DrugName("00310620530")

# Import NDC list
NDClist <- read_excel(NDCpath)
class(NDClist$NDC)
NDClist$NDC <- sprintf("%011d", as.numeric(NDClist$NDC))
# NDClist$NDC <- stringr::str_pad(NDClist$NDC, 11, side="left", pad="0")

# Apply the function
DrugName <- data.frame(Drug_Name=apply(NDClist, 1, NDC_to_DrugName))
Output <- cbind(NDClist,DrugName)

# Save the Output
write.xlsx(Output, file=Outputpath, sheetName="NDC_DrugName", row.names=FALSE)
```

## ATC to RxCUI

```{r ATC_to_RxCUI, eval=FALSE, include=TRUE}
# ATC to RxCUI --------------------------------------------------------

# Input
ATCpath <- "/Users/peilinhuang/Dropbox (UFL)/Lynn Huang/Rxmix/ATC_GLP1.xlsx"
Outputpath <- "/Users/peilinhuang/Dropbox (UFL)/Lynn Huang/Rxmix/NDC_GLP1.xlsx"

# Create function
ATC_to_RxCUI <- function(ATC){
  # Find Rxcui By ATC
  path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui.json?idtype=ATC&id=", ATC)
  RxCUI <- content(GET(url = path))$idGroup$rxnormId[[1]]
  # Get Related By RxCUI
  path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui/", RxCUI, "/related,json?tty=IN+SCD+SCDC+SBD+SBDC")
  b <- data.frame(matrix(ncol = 7, nrow = 0))
  colnames(b) <- c("rxcui", "name", "synonym", "tty", "language", "suppress", "umlscui")
  for (i in 1:5){
    max <- length(content(GET(url = path))$relatedGroup$conceptGroup[[i]]$conceptProperties)
    for (j in 1:max){
      a <- data.frame(content(GET(url = path))$relatedGroup$conceptGroup[[i]]$conceptProperties[[j]])
      b <- rbind(b,a)
    }
  }
  b <- cbind(b, "ATC"=ATC)
  b

}

# Test the function
ATC_to_RxCUI("A10BK01")

# Import ATC list
ATClist <- read_excel(ATCpath)
ATClist$ATC <- str_trim(ATClist$ATC)
ATClist <- unlist(ATClist[,1])

# Apply the function
list <- lapply(ATClist, ATC_to_NDC)
Output <- data.frame(matrix(ncol = 8, nrow = 0))
colnames(Output) <- c("rxcui", "name", "synonym", "tty", "language", "suppress", "umlscui","ATC")
for (i in 1:length(list)){
  Output <- rbind(Output, list[[i]])
}

# Save the Output
write.xlsx(Output, file=Outputpath, sheetName="sheet1", row.names=FALSE)
```

## ATC to NDC

```{r ATC_to_NDC, eval=FALSE, include=TRUE}
# ATC to NDC --------------------------------------------------------

# Input
ATCpath <- "/Users/peilinhuang/Dropbox (UFL)/Lynn Huang/R/ATC_DOAC.xlsx"
Outputpath <- "/Users/peilinhuang/Dropbox (UFL)/Lynn Huang/R/NDC_DOAC.xlsx"

# Create function
ATC_to_NDC <- function(ATC){
  # Find Rxcui By ATC
  path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui.json?idtype=ATC&id=", ATC)
  RxCUI <- content(GET(url = path))$idGroup$rxnormId[[1]]
  # Get Related By RxCUI
  path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui/", RxCUI, "/related,json?tty=IN+SCD+SCDC+SBD+SBDC")
  b <- data.frame(matrix(ncol = 7, nrow = 0))
  colnames(b) <- c("rxcui", "name", "synonym", "tty", "language", "suppress", "umlscui")
  for (i in 1:5){
    max <- length(content(GET(url = path))$relatedGroup$conceptGroup[[i]]$conceptProperties)
    for (j in 1:max){
      a <- data.frame(content(GET(url = path))$relatedGroup$conceptGroup[[i]]$conceptProperties[[j]])
      b <- rbind(b,a)
    }
  }
  b <- cbind(b, "ATC"=ATC)
  # Get NDC By RxCUI
  c <- data.frame(matrix(ncol = 4, nrow = 0))
  colnames(c) <- c("ndc", "rxcui","name","ATC")

  for (k in 1:nrow(b)) {
    rxcui <- b[k,"rxcui"]
    name <- b[k,"name"]
    ATC <- b[k,"ATC"]
    path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui/", rxcui, "/allhistoricalndcs.json")
    max <- length(content(GET(url = path))$historicalNdcConcept$historicalNdcTime[[1]]$ndcTime)
    for (l in 1:max) {
      if (is.null(content(GET(url = path))$historicalNdcConcept)) {
      } else {
        d <- data.frame("ndc"=content(GET(url = path))$historicalNdcConcept$historicalNdcTime[[1]]$ndcTime[[l]]$ndc[[1]],
                        "rxcui"=rxcui,"name"=name,"ATC"=ATC)
        c <- rbind(c,d)
      }
    }
  }
  c
}  

# Test the function
# ATC_to_NDC("A10BK01")

# Import ATC list
ATClist <- read_excel(ATCpath)
ATClist$ATC <- str_trim(ATClist$ATC)
ATClist <- unlist(ATClist[,1])

# Apply the function 
list <- lapply(ATClist, ATC_to_NDC)
Output <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(Output) <- c("ndc", "rxcui","name","ATC")
for (i in 1:length(list)){
  Output <- rbind(Output, list[[i]])
}

# Save the Output
write.xlsx(Output, file=Outputpath, sheetName="sheet1", row.names=FALSE)
```

## NDC to RxCUI

Using RxNorm API to transform pre-defined NDC code into RxCUI.

```{r NDC_to_RxCUI, eval=FALSE, include=TRUE}
# Input

NDCpath <- "/Users/lynn/Dropbox (UFL)/FL Medicaid/FL Medicaid data/HCV_OneFL_Lynn/HCV_FL_NDC.xlsx"
Outputpath <- "/Users/lynn/Dropbox (UFL)/FL Medicaid/FL Medicaid data/HCV_OneFL_Lynn/HCV_FL_NDC.xlsx"
NDClist <- read_excel(NDCpath)
NDClist$NDC <- sort(unique(NDClist$NDC))
NDClist$NDC <- str_pad(NDClist$NDC, 11, side = "left", "0")

# Create function
NDC_to_RxCUI <- function(NDC){
  # Find Rxcui By NDC
  path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui.json?idtype=NDC&id=", NDC)
  RxCUI <- content(GET(url = path))$idGroup$rxnormId[[1]]
  RxCUI <- ifelse(is.null(RxCUI),"",RxCUI)
  RxCUI
}
# Test the function
NDC_to_RxCUI("00310620530")
NDC_to_RxCUI("00004000512")
NDC_to_RxCUI("76519117002")
NDC_to_RxCUI("00024027404")



# Import NDC list
NDClist <- read_excel(NDCpath)
NDClist$NDC <- sort(unique(NDClist$NDC))
NDClist$NDC <- str_pad(NDClist$NDC, 11, side = "left", "0")
# NDClist$NDC <- stringr::str_pad(NDClist$NDC, 11, side="left", pad="0")


# Apply the function
RxCUI <- data.frame(RxCUI=apply(NDClist[,c("NDC")], 1, NDC_to_RxCUI))
Output <- cbind(NDClist,RxCUI)

# Save the Output
write.xlsx(Output, file=Outputpath, sheetName="NDC_to_RxCUI", row.names=FALSE)

```

## NDC to Strength

```{r NDC_to_Strength, eval=FALSE, include=TRUE}
# Input

NDCpath <- "C:/Users/peilinhuang/Desktop/11DigitNDC.xlsx"
Outputpath <- "C:/Users/peilinhuang/Desktop/Output.xlsx"

# Create function
NDC_to_Strength <- function(NDC){
  # Find Rxcui By NDC
  path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui.json?idtype=NDC&id=", NDC)
  RxCUI <- content(GET(url = path))$idGroup$rxnormId[[1]]
  if (is.null(RxCUI)) {
      Strength <- ""
      } else {
        # Get Strength By RxCUI
        path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui/", RxCUI, "/allProperties.json?prop=ATTRIBUTES")
        temp <- matrix(unlist(content(GET(url = path))$propConceptGroup$propConcept),nrow=3)
        colnames(temp) <-temp[2,]
        temp <- data.frame(temp)[3,]
        Strength <- temp$AVAILABLE_STRENGTH
        Strength <- ifelse(is.null(Strength),"",Strength)
      }
  return(Strength)
}
# Test the function
NDC_to_Strength("17856355504")
NDC_to_Strength("00004000512")
NDC_to_Strength("00093613832")


# Import NDC list
NDClist <- read_excel(NDCpath)
NDClist$NDC <- str_pad(NDClist$NDC, 11, side = "left", "0")
# NDClist$NDC <- stringr::str_pad(NDClist$NDC, 11, side="left", pad="0")

# Apply the function
Strength <- data.frame(Strength=apply(NDClist, 1, NDC_to_Strength))
Output <- cbind(NDClist,Strength)

# Save the Output
write.xlsx(Output, file=Outputpath, sheetName="sheet1", row.names=FALSE)

```

## NDC to Porperties

```{r NDC_to_Porperties, eval=FALSE, include=TRUE}

# Input

NDCpath     <- "/Users/peilinhuang/Dropbox (UFL)/Lynn Huang/Rxmix/11DigitNDC/11DigitNDC.xlsx"
Outputpath  <- "C:/Users/peilinhuang/Desktop/Output.xlsx"
Property    <- sort(c("NDC","RxCUI","TTY", "AVAILABLE_STRENGTH","STRENGTH","RxNorm Name","QUANTITY"))
Property
# Create function
NDC_to_Porperties <- function(NDC,Property){
  # Find Rxcui By NDC
  path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui.json?idtype=NDC&id=", NDC)
  RxCUI <- content(GET(url = path))$idGroup$rxnormId[[1]]
  if (is.null(RxCUI)) {
      temp3 <- matrix(, nrow=1, ncol=length(Property))
      } else {
        # Get Properties By RxCUI
        path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui/", RxCUI,"/allProperties.json?prop=NAMES+ATTRIBUTES")
        temp <- matrix(unlist(content(GET(url = path))$propConceptGroup$propConcept),nrow=3)
        temp <- t(cbind(c("NDC",NDC),c("RxCUI",RxCUI),temp[2:3,]))
        temp2 <- matrix(Property, nrow = length(Property), byrow = TRUE)
        temp3 <- t(merge(x=temp2,y=temp,by="V1",all.x=TRUE))[2,]
      }
  return(temp3)
}

# Test the function

a <- NDC_to_Porperties("00310620530",Property)
b <- NDC_to_Porperties("00004000512",Property)
c <- NDC_to_Porperties("00093613832",Property)
d <- NDC_to_Porperties("00002144561",Property)
NDC_to_Porperties("00310620530",Property)

# Import NDC list
NDClist <- read_excel(NDCpath)
NDClist$NDC <- str_pad(NDClist$NDC, 11, side = "left", "0")
NDClist <- NDClist[1:10,]

# Apply the function
Property_list<- lapply(unlist(NDClist), function(x)NDC_to_Porperties(x,Property))

Output <- matrix(Property,nrow=1, ncol =length(Property))

for (i in 1:nrow(NDClist)){
  Output <- rbind(Output,matrix(Property_list[[i]],nrow=1, ncol =length(Property)))
}
Output <- data.frame(Output)[-1,]
colnames(Output) <- Property

# Save the Output
write.xlsx(Output, file=Outputpath, sheetName="sheet1", row.names=FALSE)
```

```{r}
# find RXCUI BY STRING
# Hydrocodone

# Input

NDCpath     <- "/Users/peilinhuang/Dropbox (UFL)/Lynn Huang/Rxmix/11DigitNDC/11DigitNDC.xlsx"
Outputpath  <- "C:/Users/peilinhuang/Desktop/Output.xlsx"
Property    <- sort(c("NDC","RxCUI","TTY", "AVAILABLE_STRENGTH","STRENGTH","RxNorm Name","QUANTITY"))

path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui.json?name=hydrocodone&allsrc=0&search=2")
df <- content(GET(url = path))
RXCUI <- 

# Test the function
NDC_to_Porperties <- function(NDC,Property){
  # Find Rxcui By NDC
  path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui.json?name=",DrugName,"&allsrc=0&search=2", )
  RxCUI <- content(GET(url = path))$idGroup$rxnormId[[1]]
  if (is.null(RxCUI)) {
      temp3 <- matrix(, nrow=1, ncol=length(Property))
      } else {
        # Get Properties By RxCUI
        path <- str_c("https://rxnav.nlm.nih.gov/REST/rxcui/", RxCUI,"/allProperties.json?prop=NAMES+ATTRIBUTES")
        temp <- matrix(unlist(content(GET(url = path))$propConceptGroup$propConcept),nrow=3)
        temp <- t(cbind(c("NDC",NDC),c("RxCUI",RxCUI),temp[2:3,]))
        temp2 <- matrix(Property, nrow = length(Property), byrow = TRUE)
        temp3 <- t(merge(x=temp2,y=temp,by="V1",all.x=TRUE))[2,]
      }
  return(temp3)
}

```
